\Huge \bf \sc Calibration of an inertial measurement unit

### 2013-07-30 latex/paper_sensores/paper.tex
a ver si queda

\begin{abstract}
%\boldmath
This paper presents a way to calibrate different inertial measurement sensors. In particular we present the calibration of an accelerometer and a gyroscope using nonlinear least squares. A model of the sensors is presented based on the main errors that MEMS devices present, a calibration method is proposed for the static parameters of the model. Finally a temperature adjust is made.  
% -cambio-
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

The calibration proposed in this paper is tested on a 3-axis accelerometer ADXL345 of Analog Devices and a 3-axis gyroscope ITG-3200  of InvenSense, yet the model developed and the methodology can be used in other devices because the model is based on common characteristics of MEMS sensors. The sensors named before are included in a Inertial Measurement Unit(IMU): The Mongoose (figure \refp{fig:mongoose}). \\

@@

In the datasheets of the considered sensors(\cite{bib:acc_data} and \cite{bib:gyro_data}) the effect of nonlinearities negligible since it represents the $\pm 0.5.\%$  of the full scale for the accelerometer and the $\pm 0.2\%$ for the gyroscope. Therefore we will not consider this effect and after the calibration we must verify that the error due to the nonlinear response is negligible.\\
%TODO Al final hacer las cuentas para ver que es despreciable.

In Figure \refp{fig:noise} we can observe that the electrical noise of the devices is typically 2 bits, this error (according to the datasheets \cite{bib:acc_data} and \cite{bib:gyro_data}) correspond to $7.8 mg$ and $0.14 ^{\circ}/s $. The error is not biased, thus we not need to consider it, because several samples are used in any real application and the effect vanishes.\\

@@

\subsection{Accelerometer}
Due to construction issues the sensitivity axis of the device are generally not orthogonal. Lets define a platform system as shown in Figure \refp{fig:mongoose}. Let $\mathbf{a}^a$ be the true acceleration (meaning the theoretical acceleration measured in $ms^{-2}$) expressed in the sensitivity axis of the accelerometer and let $\mathbf{a}^p$ be the theoretical true acceleration (also measured in $ms^{-2}$) expressed in the platform system. We model this fact by the following linear transformation between the two acceleration vectors:
\begin{equation}
\mathbf{a}^p = \mathbf{T}_a^p\mathbf{a}^a, \; \mathbf{T}_a^p = \left(\begin{array}{ccc}
1 & - \alpha_{yz} & \alpha_{zy}\\
\alpha_{xz} & 1 & - \alpha_{zx}\\
-\alpha_{xy} & \alpha_{yx} & 1
\end{array}\right)
\label{eq:nonorto}
\end{equation}
In equation \refp{eq:nonorto}, scalars $\alpha_{ij}$ represents the rotation of the $i-th$ sensitivity axis of the accelerometer over the $j-th$ axis of the platform system. In Figure \refp{fig:nonorto} this relation can be observed graphically. Since those errors are due to the manufacturing process we will assume that they will remain constant long enough.

@@

As it was stated before, we are considering a linear model between the acceleration measured in the sensitivity axis and the true acceleration in the same system. Thus we have:
\begin{equation}
\tilde{\mathbf{a}}^a = \mathbf{K}_a\mathbf{a}^a + \mathbf{b}_a
\label{eq:modlineal}
\end{equation}
In equation \refp{eq:modlineal}, $\tilde{\mathbf{a}}^a$ is the measure obtained by the accelerometer (measured in bits), $\mathbf{K}_a$ is a diagonal matrix that represents the sensitivity of each axis and $\mathbf{b}_a$ is a vector that represent the bias of each axis. Those parameters may variate as the temperature changes. However, in this stage we are going to consider them constant. Using equations \refp{eq:nonorto} and \refp{eq:modlineal} we can establish a model of the accelerometer:

@@

It is important to consider that the function that we want to minimize will generally have more than one local minimum. Thus, we can only assure that the optimum found by algorithms is a local optimum. To ensure that the solution obtained is the desired one, we need to choose carefully the seed for the algorithm. Typically we will choose a vector $\boldsymbol{\theta}_0$ that is close to the values of the unknown parameters. Even if we do not know the exact value of the model's parameters, the datasheet of the different sensors can give us a very good clue about those values.\\

@@

The 27 measures made were the following:
\begin{itemize}
\item Starting position: $x$ axis pointing ``down'': 10 measures were made combining two rotations. First, rotating around the $z$ axis $\theta =0^\circ, 10^\circ, 20^\circ, 30^\circ, 45^\circ$. Then, for each angle defined(except for $\theta = 0^\circ$), two measures were made rotating around the $x$ axis $\psi = 0^\circ, -45^\circ$. This set of measures can be expressed in the following way:
\begin{scriptsize}
\begin{equation}
\mathbf{a}^p = \left(\begin{array}{ccc}
1 & 0 & 0\\
0 & \cos \psi & \sin \psi \\
0 & -\sin \psi & \cos \psi
\end{array}\right)\left(\begin{array}{ccc}
\cos \theta & \sin \theta & 0\\
-\sin \theta & \cos \theta & 0\\
0 & 0 & 1
\end{array}\right)\left(\begin{array}{c}
-g\\
0\\
0
\end{array}\right)
\label{eq:acc_x}
\end{equation}
\end{scriptsize}

\item Starting position: $y$ axis pointing ``up'': 10 measures were made combining two rotations. First, rotating around the $x$ axis $\psi =0^\circ, 10^\circ, 20^\circ, 30^\circ, 45^\circ$. Then, for each angle defined(except for $\theta = 0^\circ$), two measures were made rotating around the $y$ axis $\varphi = 0^\circ, 45^\circ$.

@@

\item Starting position: $z$ axis pointing ``up'': 10 measures were made combining two rotations. First, rotating around the $y$ axis: $\phi =0^\circ, -10^\circ, -20^\circ, -30^\circ, -45^\circ$. Then, for each angle defined(except for $\phi = 0^\circ$), two measures were made rotating around the $z$ axis $\theta = 0^\circ, 45^\circ$.   \end{itemize}

@@

\begin{figure}[h!]
	\subfloat[Accelerometer measurement in\newline z-axis while heating up the sensor.]{\label{fig:resultado_temp} 
			\includegraphics[width=.5\columnwidth]{./pics_paper/resultado_temp.pdf}}
	\subfloat[Accelerometer measurements and temperature varying from $35^oC$ to $48^oC$.]{\label{fig:bajada}
		\includegraphics[width=.5\columnwidth]{./pics_paper/bajada.pdf}}
	\caption{\textbf{Temperature results -} Accelerometer measurements while temperature is modified.}
	\label{fig:temp_res}
\end{figure}

>>>>

\begin{abstract}
%\boldmath
This paper presents a fast and low cost way to calibrate different inertial measurement sensors. In particular we present the calibration of an accelerometer and a gyroscope using nonlinear least squares. A model of the sensors is presented based on the main errors that MEMS devices present, a calibration method is proposed for the static parameters of the model. Finally a temperature adjust is made.  
% -cambio-
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

The calibration proposed in this work is tested on a 3-axis accelerometer \emph{ADXL345} of \emph{Analog Devices} and a 3-axis gyroscope \emph{ITG-3200} of \emph{InvenSense}, yet the model developed and the methodology can be used in other devices because the model is based on common characteristics of MEMS sensors. The sensors named before are included in a Inertial Measurement Unit(IMU): The Mongoose (figure \refp{fig:mongoose}). \\

The motivation for this calibration is to have enough precision to estimate the state of an Unmanned Aerial Vehicle (UAV) with quadrotor architecture.\\

@@

According to the datasheets of the devices (\cite{bib:acc_data, bib:gyro_data}), effect of nonlinearities represent the $\pm 0.5.\%$  of the full scale for the accelerometer and the $\pm 0.2\%$ for the gyroscope. Therefore, this effect is not considered and after the calibration, it must be verified.\\
%TODO Al final hacer las cuentas para ver que es despreciable.

In Figure \refp{fig:noise} we can observe that the electrical noise of the devices is typically 3 LSB (Least Significant Bit) peak to peak, which (according to the datasheets \cite{bib:acc_data, bib:gyro_data}) correspond to $11.7 mg$ and $0.21 ^{\circ}/s $. The error is not biased, thus, we not need to consider it, because many samples are used in the procedure and the effect vanishes.\\

@@

\subsection{Accelerometer}
Due to construction issues the sensitivity axis of the device are generally not orthogonal. Lets define a platform system as shown in Figure \refp{fig:mongoose}. Let $\mathbf{a}^a$ be the true acceleration (meaning the theoretical acceleration measured in $ms^{-2}$) expressed in the sensitivity axis of the accelerometer and let $\mathbf{a}^p$ be the theoretical true acceleration (also measured in $ms^{-2}$) expressed in the platform system. The two acceleration vectors are related according to equation \refp{eq:nonorto}:
\begin{equation}
\mathbf{a}^p = \mathbf{T}_a^p\mathbf{a}^a, \; \mathbf{T}_a^p = \left(\begin{array}{ccc}
1 & - \alpha_{yz} & \alpha_{zy}\\
\alpha_{xz} & 1 & - \alpha_{zx}\\
-\alpha_{xy} & \alpha_{yx} & 1
\end{array}\right)
\label{eq:nonorto}
\end{equation}
In equation \refp{eq:nonorto}, scalars $\alpha_{ij}$ represents the rotation of the $i-th$ sensitivity axis of the accelerometer over the $j-th$ axis of the platform system. In Figure \refp{fig:nonorto} this relation can be observed graphically. Since those errors are due to the manufacturing process we will assume that they will remain constant long enough.

@@

As it was stated before, a linear model between the acceleration measured in the sensitivity axis and the true acceleration in the same system is considered. Thus we have:
\begin{equation}
\tilde{\mathbf{a}}^a = \mathbf{K}_a\mathbf{a}^a + \mathbf{b}_a
\label{eq:modlineal}
\end{equation}
In equation \refp{eq:modlineal}, $\tilde{\mathbf{a}}^a$ is the measure obtained by the accelerometer (measured in bits), $\mathbf{K}_a$ is a diagonal matrix that represents the sensitivity of each axis and $\mathbf{b}_a$ is a vector that represent the bias of each axis. Those parameters may variate as the temperature changes. However, constant temperature is considered. Using equations \refp{eq:nonorto} and \refp{eq:modlineal} we can establish a model of the accelerometer:

@@

It is important to consider that the minimize function will generally have more than one local minimum. To ensure that the solution obtained is the desired one, the seed for the algorithm must be carefully chosen. Typically we will choose a vector $\boldsymbol{\theta}_0$ that is close to the values of the unknown parameters. Even if we do not know the exact value of the model's parameters, the datasheet of the different sensors can give us a very good clue about those values.\\

@@

The 27 measures made were the following:
\begin{itemize}
\item Starting position: $x$ axis pointing ``down'': 9 measures were made combining two rotations. First, rotating around the $z$ axis $\theta =0^\circ, 10^\circ, 20^\circ, 30^\circ, 45^\circ$. Then, for each angle defined(except for $\theta = 0^\circ$), two measures were made rotating around the $x$ axis $\psi = 0^\circ, -45^\circ$. This set of measures can be expressed in the following way:
\begin{scriptsize}
\begin{equation}
\mathbf{a}^p = \left(\begin{array}{ccc}
1 & 0 & 0\\
0 & \cos \psi & \sin \psi \\
0 & -\sin \psi & \cos \psi
\end{array}\right)\left(\begin{array}{ccc}
\cos \theta & \sin \theta & 0\\
-\sin \theta & \cos \theta & 0\\
0 & 0 & 1
\end{array}\right)\left(\begin{array}{c}
-g\\
0\\
0
\end{array}\right)
\label{eq:acc_x}
\end{equation}
\end{scriptsize}

\item Starting position: $y$ axis pointing ``up'': 9 measures were made combining two rotations. First, rotating around the $x$ axis $\psi =0^\circ, 10^\circ, 20^\circ, 30^\circ, 45^\circ$. Then, for each angle defined(except for $\theta = 0^\circ$), two measures were made rotating around the $y$ axis $\varphi = 0^\circ, 45^\circ$.

@@

\item Starting position: $z$ axis pointing ``up'': 9 measures were made combining two rotations. First, rotating around the $y$ axis: $\phi =0^\circ, -10^\circ, -20^\circ, -30^\circ, -45^\circ$. Then, for each angle defined(except for $\phi = 0^\circ$), two measures were made rotating around the $z$ axis $\theta = 0^\circ, 45^\circ$.   \end{itemize}

@@

\begin{figure}[h!]
	\subfloat[Accelerometer measurement in\newline z-axis vs. temperature while heating\newline up the sensor.]{\label{fig:resultado_temp} 
			\includegraphics[width=.5\columnwidth]{./pics_paper/resultado_temp.png}}
	\subfloat[Accelerometer measurements and temperature (T) vs. time while T varying from $35^oC$ to $48^oC$.]{\label{fig:bajada}
		\includegraphics[width=.5\columnwidth]{./pics_paper/bajada.png}}
	\caption{\textbf{Temperature results -} Accelerometer measurements while temperature is modified.}
	\label{fig:temp_res}
\end{figure}
